\chapter{Interview guide (stage 3)}
\label{appendix-interview-guide-stage-3}

\newpage
	
\hrulefill\par

\begin{center}
\textbf{Interview guide (English)}
\end{center}

\begin{enumerate}
	\item Explain purpose of the research, summarise consent form guidelines and ask for doubts.
	\item Introductory questions:
		\begin{enumerate}
			\item Could you tell me about when and why your interest in Drupal began?
			\item How do you make your living? What is your main role in the projects in which you use Drupal?
			\item Are\slash were you involved in other FLOSS communities? Could you compare your personal experiences of the Drupal community and the others? 
			\item What does Drupal mean to you?
		\end{enumerate}
	
	\item Core initiatives - Twig in Core:
		\begin{enumerate}
			\item Introductory questions about the initiative:
				\begin{enumerate}
					\item Could you explain to me what core initiatives consist of? What about, more specifically, Twig in Core?
					\item Could you explain to me how the Twig in Core initiative started?
					\item How did you start/get involved in the development of Twig in Core? When did you start?
					\item Could you explain to me how the initiative became official? Was it appointed by Dries?
					\item Did the initiative have to pass the core gates? Could you explain that process to me? If not, do you think this facilitates the process for non-official initiatives?
					\item Do you think an initiative like this would have been possible 7 or 8 years ago? For example, in the transition between previous versions of Drupal (5 to 6, or 6 to 7).
				\end{enumerate}		
			\item Day-to-day
				\begin{enumerate}
					\item How could people participate in Twig in Core? How many people participated in total? Could you describe how the participation changed over time? For example: was it sporadic? Did it grow or decrease over time?
					\item Were there different periods? Could you explain to me how you managed to get Twig into core? Could you tell me about the ``sword of consensus"?
					\item Could you explain to me what the day-to-day was like in Twig in Core?
					\item How were the decisions taken in the initiative once Twig was in core (e.g. simply vs classy)? For example, adding or not new patches, creating or not a new release, or creating or not a new feature. Was there rotation in the leadership? How were decisions made around the roadmap? Could you explain to me the ``consensus banana" story? What happened?
					\item Did you organise any other initiatives around Twig in Core (e.g. events, working groups, etc.) to strengthen the initiative? When and why did this happen?
					\item How did the idea of Frontend United started? Could you explain to me what the event was about? Do you think events like these were relevant to strengthen the initiative? Were there any similar events before? Do you think frontend developers had more of a voice over the years in the Drupal community?
					\item Do you think it was harder to get an initiative focussed on frontend into core (rather than backend)? Why do you think it was easier/harder? How useful were the Core Gates for this? Do you think themers' voices are less heard?
				\end{enumerate}			
		\item Leadership and empowerment			
			\begin{enumerate}	
					\item Were there any ``newbie" contributors in the initiative? What was their experience like?
					\item Could you explain to me what is involved by being an initiative leader. Were there any initiative leaders in Twig in Core? How were they/you appointed? 
					\item Could you explain to me when the idea of having official initiative leaders started? Why do you think this happened?
					\item Do you have the impression that the processes related to the development of core are becoming more standardised and/or formalised over time? What do you think could be the consequences?
			\end{enumerate}	

		\end{enumerate}
	
	\item Are there any issues that you expected me to ask about, or that you think I should know?
	\item Closing up. Jot down impressions.
\end{enumerate}